---
title: "Weight descriptions for collaborative ensembles"
author: "Nicholas G Reich"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
## weight checking and visualization
library(ggplot2)
library(directlabels)
library(dplyr)

theme_set(theme_minimal())
```

## Constant weights

```{r}
constant_weights <- read.csv("../weights/constant-weights.csv") %>%
    mutate(component_model_id = reorder(component_model_id, weight))
ggplot(constant_weights, aes(x=as.numeric(season), y=weight, color=component_model_id)) + 
    geom_line() + 
    geom_dl(aes(label = component_model_id), method="last.points", cex=.8) +
    scale_x_continuous(
        limits=c(1, 10), breaks=1:7,
        labels = paste0(2010:2016, "/", 2011:2017),
        name=NULL) + 
    scale_colour_discrete(guide = 'none') +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    ggtitle("Constant model weights for each left-out season")

ggplot(constant_weights, aes(x=season, fill=weight, y=component_model_id)) + 
    geom_tile() + ylab(NULL) + xlab(NULL) +
    geom_text(aes(label=round(weight, 2))) +
    scale_fill_gradient(low = "white", high="dodgerblue4", limits=c(0,1)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    ggtitle("Constant model weights by season")
```

```{r}
constant_weights %>%
    group_by(season) %>%
    summarize(total_weights = sum(weight))
```

## Target-type weights

```{r}
target_type_weights <- read.csv("../weights/target-type-based-weights.csv") %>%
    filter(target == "1 wk ahead" | target=="Season onset") %>%
    mutate(component_model_id = reorder(component_model_id, weight))

ggplot(target_type_weights, aes(x=as.numeric(season), y=weight, color=component_model_id)) + 
    geom_line() + 
    geom_dl(aes(label = component_model_id), method="last.points", cex=.8) +
    scale_x_continuous(
        limits=c(1, 10), breaks=1:7,
        labels = paste0(2010:2016, "/", 2011:2017),
        name=NULL) + 
    scale_colour_discrete(guide = 'none') +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    facet_grid(target~.) +
    ggtitle("Target-type model weights for each left-out season")

avg_target_type_weights <- target_type_weights %>% 
    group_by(target, component_model_id) %>% 
    summarize(weight = mean(weight)) %>% ungroup()
    
ggplot(avg_target_type_weights, aes(x=target, fill=weight, y=component_model_id)) + 
    geom_tile() + ylab(NULL) + xlab(NULL) +
    geom_text(aes(label=round(weight, 2))) +
    scale_fill_gradient(low = "white", high="dodgerblue4", limits=c(0,1)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    ggtitle("Average LOSO model weights by target")


target_type_weights %>%
    group_by(season, target) %>%
    summarize(total_weights = sum(weight))
```

## Target-specific weights

```{r}
target_weights <- read.csv("../weights/target-based-weights.csv")
avg_target_weights <- target_weights %>% 
    group_by(target, component_model_id) %>% 
    summarize(weight = mean(weight)) %>% ungroup() %>%
    mutate(component_model_id = reorder(component_model_id, weight))
ggplot(avg_target_weights, aes(x=target, fill=weight, y=component_model_id)) + 
    geom_tile() + ylab(NULL) + xlab(NULL) +
    geom_text(aes(label=round(weight, 2))) +
    scale_fill_gradient(low = "white", high="dodgerblue4", limits=c(0,1)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    ggtitle("Average LOSO model weights by target")
```


Do all weights sum to 1 within a season and target?
```{r echo=TRUE}
tot_target_weights <- target_weights %>%
    group_by(season, target) %>%
    summarize(total_weights = sum(weight)) 
all.equal(
    tot_target_weights$total_weights, 
    rep(1, nrow(tot_target_weights))
    )
```



## Target and region weights

```{r, fig.height=10, fig.width=10}
target_region_weights <- read.csv("../weights/target-and-region-based-weights.csv")
avg_target_region_weights <- target_region_weights %>% 
    group_by(target, location, component_model_id) %>% 
    summarize(weight = mean(weight)) %>% ungroup() %>%
    mutate(component_model_id = reorder(component_model_id, weight),
        weight_text = ifelse(weight<0.005, NA, weight))
ggplot(avg_target_region_weights, 
    aes(x=target, fill=weight, y=component_model_id)) + 
    geom_tile() + ylab(NULL) + xlab(NULL) +
    geom_text(aes(label=round(weight_text, 2)), size=2) +
    scale_fill_gradient(low = "white", high="dodgerblue4", limits=c(0,1)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    facet_wrap("location") + 
    ggtitle("Average LOSO model weights by target and region")
```

Do all weights sum to 1 within a season and target?
```{r echo=TRUE}
tot_target_region_weights <- target_region_weights %>%
    group_by(season, target, location) %>%
    summarize(total_weights = sum(weight))
all.equal(
    tot_target_region_weights$total_weights, 
    rep(1, nrow(tot_target_region_weights))
    )
